---
title: "Pink Floyd & Stone Temple Pilots Albums"
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(knitr)
library(plotly)

albums <- read_csv("albums.csv", show_col_types = FALSE)

similarity_titles <- c(
  "The Division Bell",
  "Purple",
  "The Piper at the Gates of Dawn",
  "Perdida"
)

render_gallery <- function(data) {
  cards <- glue::glue(
    '<div class="album-card">',
    '<img src="{data$cover_url}" alt="{data$title} by {data$artist}" class="album-cover" loading="lazy" onerror="this.onerror=null;this.src=\'https://placehold.co/250x250/222/fff?text=Cover+Unavailable\';" />',
    '<div class="album-popup">',
    '<div class="popup-title">{data$title}</div>',
    '<div class="popup-artist">{data$artist}</div>',
    '<span class="popup-date">Released: {data$release_date}</span>',
    '</div>',
    '</div>'
  )
  cat(paste0('<div class="album-gallery">\n', paste(cards, collapse = "\n"), '\n</div>\n'))
}
```

This site compares the studio album discographies of **Pink Floyd** and **Stone Temple Pilots** using release date data from MusicBrainz.

## Discography Timeline & Popularity

The interactive chart below plots each studio album by release date and user rating from MusicBrainz. Point size indicates the number of user votes. Hover over any point to inspect an album.

```{r}
#| label: timeline-popularity-plot
#| warning: false

p <- albums |>
  mutate(
    date = as.Date(release_date),
    tooltip = paste0(
      "<b>", title, "</b><br>",
      artist, "<br>",
      "Released: ", release_date, "<br>",
      "Rating: ", rating, " / 5 (", votes, " votes)"
    )
  ) |>
  ggplot(aes(x = date, y = rating, color = artist, text = tooltip)) +
  geom_point(aes(size = votes), alpha = 0.85) +
  scale_y_continuous(limits = c(3, 5), breaks = seq(3, 5, 0.5)) +
  scale_color_manual(values = c("Pink Floyd" = "#e76f51", "Stone Temple Pilots" = "#2a9d8f")) +
  theme_minimal(base_size = 12) +
  labs(
    title = "Album Ratings & Releases (1967–2020)",
    x = "Release Date",
    y = "Rating (out of 5)",
    color = "Artist",
    size = "Votes"
  )

ggplotly(p, tooltip = "text")
```

## Album Similarity

Albums released closest in time represent the bands' most similar era, while those furthest apart are the least similar. Hover over any album cover to see details.

### Most Similar Albums (70 Days Apart)

- **Pink Floyd:** *The Division Bell* (March 28, 1994)
- **Stone Temple Pilots:** *Purple* (June 6, 1994)

```{r}
#| label: most-similar-albums
#| results: asis

albums |>
  filter(
    (artist == "Pink Floyd" & title == "The Division Bell") |
    (artist == "Stone Temple Pilots" & title == "Purple")
  ) |>
  render_gallery()
```

### Least Similar Albums (52+ Years Apart)

- **Pink Floyd:** *The Piper at the Gates of Dawn* (July 7, 1967)
- **Stone Temple Pilots:** *Perdida* (February 7, 2020)

```{r}
#| label: least-similar-albums
#| results: asis

albums |>
  filter(
    (artist == "Pink Floyd" & title == "The Piper at the Gates of Dawn") |
    (artist == "Stone Temple Pilots" & title == "Perdida")
  ) |>
  render_gallery()
```

## Other Albums by Band

Hover your cursor over any album cover to see its details.

### Pink Floyd

```{r}
#| label: pink-floyd-albums
#| results: asis

albums |>
  filter(artist == "Pink Floyd", !title %in% similarity_titles) |>
  render_gallery()
```

### Stone Temple Pilots

```{r}
#| label: stp-albums
#| results: asis

albums |>
  filter(artist == "Stone Temple Pilots", !title %in% similarity_titles) |>
  render_gallery()
```

