---
title: "Pink Floyd & Stone Temple Pilots Albums"
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(knitr)

albums <- read_csv("albums.csv", show_col_types = FALSE)

similarity_titles <- c(
  "The Division Bell",
  "Purple",
  "The Piper at the Gates of Dawn",
  "Perdida"
)

render_gallery <- function(data) {
  cards <- glue::glue(
    '<div class="album-card">',
    '<img src="{data$cover_url}" alt="{data$title} by {data$artist}" class="album-cover" loading="lazy" onerror="this.onerror=null;this.src=\'https://placehold.co/250x250/222/fff?text=Cover+Unavailable\';" />',
    '<div class="album-popup">',
    '<div class="popup-title">{data$title}</div>',
    '<div class="popup-artist">{data$artist}</div>',
    '<span class="popup-date">Released: {data$release_date}</span>',
    '</div>',
    '</div>'
  )
  cat(paste0('<div class="album-gallery">\n', paste(cards, collapse = "\n"), '\n</div>\n'))
}
```

This site compares the studio album discographies of **Pink Floyd** and **Stone Temple Pilots** using release date data from MusicBrainz.

## Album Similarity

Albums released closest in time represent the bands' most similar era, while those furthest apart are the least similar. Hover over any album cover to see details.

### Most Similar Albums (70 Days Apart)

- **Pink Floyd:** *The Division Bell* (March 28, 1994)
- **Stone Temple Pilots:** *Purple* (June 6, 1994)

```{r}
#| label: most-similar-albums
#| results: asis

albums |>
  filter(
    (artist == "Pink Floyd" & title == "The Division Bell") |
    (artist == "Stone Temple Pilots" & title == "Purple")
  ) |>
  render_gallery()
```

### Least Similar Albums (52+ Years Apart)

- **Pink Floyd:** *The Piper at the Gates of Dawn* (July 7, 1967)
- **Stone Temple Pilots:** *Perdida* (February 7, 2020)

```{r}
#| label: least-similar-albums
#| results: asis

albums |>
  filter(
    (artist == "Pink Floyd" & title == "The Piper at the Gates of Dawn") |
    (artist == "Stone Temple Pilots" & title == "Perdida")
  ) |>
  render_gallery()
```

## Other Albums by Band

Hover your cursor over any album cover to see its details.

### Pink Floyd

```{r}
#| label: pink-floyd-albums
#| results: asis

albums |>
  filter(artist == "Pink Floyd", !title %in% similarity_titles) |>
  render_gallery()
```

### Stone Temple Pilots

```{r}
#| label: stp-albums
#| results: asis

albums |>
  filter(artist == "Stone Temple Pilots", !title %in% similarity_titles) |>
  render_gallery()
```

